**The Heart can Lie: The Role of Interoception and Theory of Mind in Deception**

### 2023-06-05 paper/manuscript.Rmd
minor

  - name          : "Dominique Makowski"
    affiliation   : "1, 2"
    corresponding : yes    # Define only one corresponding author
    address       : "Pevensey 1, University of Sussex, Brighton, UK"
    email         : "D.Makowski@sussex.ac.uk"
    orcid         : 0000-0001-5375-9967
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - "Conceptualization"
      - "Data curation"
      - "Formal Analysis"
      - "Funding acquisition"
      - "Investigation"
      - "Methodology"
      - "Project administration"
      - "Resources"
      - "Software"
      - "Supervision"
      - "Validation"
      - "Visualization"
      - "Writing – original draft"
  - name          : "Zen J. Lau"  # lauzenjuen@gmail.com
    affiliation   : " 2"
  - name          : "Tam Pham"  # phamttam17@gmail.com
    affiliation   : " 2"
  - name          : "An Shu Te" # anshu.te@ntu.edu.sg
    affiliation   : "2"
    orcid         : 0000-0001-8008-2824
    role:
      - "Project administration"
      - "Resources"
      - "Investigation"
      - "Writing – original draft"
  - name          : "Stephanie Kirk"  # stephanie.morgank@gmail.com
    affiliation   : "2"
    orcid         : 0000-0002-9312-5552
    role:
      - "Project administration"
      - "Resources"
      - "Writing – original draft"
  - name          : "Claudia Liauw"  # YLIAUW001@e.ntu.edu.sg
    affiliation   : "2"
    role:
      - "Data curation"
      - "Formal Analysis"
  - name          : "S.H. Annabel Chen"  # AnnabelChen@ntu.edu.sg
    affiliation   : "2, 3, 4, 5"
    orcid         : 0000-0002-1540-5516
    role:
      - "Project administration"
      - "Supervision"
      - "Writing – review & editing"

affiliation:
  - id            : "1"
    institution   : "School of Psychology, University of Sussex, UK"
  - id            : "2"
    institution   : "School of Social Sciences, Nanyang Technological University, Singapore"
  - id            : "3"
    institution   : "Centre for Research and Development in Learning, Nanyang Technological University, Singapore"
  - id            : "4"
    institution   : "Lee Kong Chian School of Medicine, Nanyang Technological University, Singapore"
  - id            : "5"
    institution   : "National Institute of Education, Nanyang Technological University, Singapore"
abstract: |
  While a large part of the deception literature focuses on lying detection, the factors contributing to one's ability to lie remain unclear. The present study examined the contribution of Theory of Mind (ToM) and interoception on our ability to lie using a directed lie paradigm with two conditions ("Interrogation" and "Polygraph"), designed to enhance each of the two mechanisms. Using a mixed-design Bayesian analysis, we show that various facets of interoceptive abilities are positively related to the self-rated confidence in one's own lies, especially when under the belief that bodily signals are being monitored (i.e., in the "Polygraph" condition). Beyond providing evidence for the role of the body in lying and raising interesting questions for deception science, these results carry practical implications for criminology and lie detection protocols.
output:
  papaja::apa6_pdf:
    keep_tex: yes
keywords: "Deception; Interoception; Theory of Mind; Poygraph; Lying Ability"
wordcount: '3720'
bibliography: ["references.bib"]
floatsintext: yes
figsintext: yes
figurelist: no
tablelist: no
footnotelist: no
linenumbers: yes
mask: yes
draft: no
documentclass: apa6
classoption: man
header-includes:
- \usepackage[labelfont=bf, font={color=gray,small}]{caption}
- \usepackage{float}
- \usepackage[document]{ragged2e}
csl: utils/apa.csl
---

>>>>

  - name          : "Dominique Makowski"
    affiliation   : "1, 2"
    corresponding : yes    # Define only one corresponding author
    address       : "Pevensey 1, University of Sussex, Brighton, UK"
    email         : "D.Makowski@sussex.ac.uk"
    orcid         : 0000-0001-5375-9967
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - "Conceptualization"
      - "Data curation"
      - "Formal Analysis"
      - "Funding acquisition"
      - "Investigation"
      - "Methodology"
      - "Project administration"
      - "Resources"
      - "Software"
      - "Supervision"
      - "Validation"
      - "Visualization"
      - "Writing – original draft"
  - name          : "Zen J. Lau"  # lauzenjuen@gmail.com
    affiliation   : " 2"
    role:
      - "Data curation"
      - "Project administration"
      - "Resources"
      - "Investigation"
  - name          : "Tam Pham"  # phamttam17@gmail.com
    affiliation   : " 2"
    role:
      - "Data curation"
      - "Project administration"
      - "Resources"
      - "Investigation"
  - name          : "An Shu Te" # anshu.te@ntu.edu.sg
    affiliation   : "2"
    orcid         : 0000-0001-8008-2824
    role:
      - "Project administration"
      - "Resources"
      - "Investigation"
      - "Writing – original draft"
  - name          : "Stephanie Kirk"  # stephanie.morgank@gmail.com
    affiliation   : "2"
    orcid         : 0000-0002-9312-5552
    role:
      - "Project administration"
      - "Resources"
      - "Writing – original draft"
  - name          : "Claudia Liauw"  # YLIAUW001@e.ntu.edu.sg
    affiliation   : "2"
    role:
      - "Data curation"
      - "Formal Analysis"
  - name          : "S.H. Annabel Chen"  # AnnabelChen@ntu.edu.sg
    affiliation   : "2, 3, 4, 5"
    orcid         : 0000-0002-1540-5516
    role:
      - "Project administration"
      - "Supervision"
      - "Writing – review & editing"

affiliation:
  - id            : "1"
    institution   : "School of Psychology, University of Sussex, UK"
  - id            : "2"
    institution   : "School of Social Sciences, Nanyang Technological University, Singapore"
  - id            : "3"
    institution   : "Centre for Research and Development in Learning, Nanyang Technological University, Singapore"
  - id            : "4"
    institution   : "Lee Kong Chian School of Medicine, Nanyang Technological University, Singapore"
  - id            : "5"
    institution   : "National Institute of Education, Nanyang Technological University, Singapore"
abstract: |
  While a large part of the deception literature focuses on lying detection, the factors contributing to one's ability to lie remain unclear. The present study examined the contribution of Theory of Mind (ToM) and interoception on our ability to lie using a directed lie paradigm with two conditions ("Interrogation" and "Polygraph"), designed to enhance each of the two mechanisms. Using a mixed-design Bayesian analysis, we show that various facets of interoceptive abilities are positively related to the self-rated confidence in one's own lies, especially when under the belief that bodily signals are being monitored (i.e., in the "Polygraph" condition). Beyond providing evidence for the role of the body in lying and raising interesting questions for deception science, these results carry practical implications for criminology and lie detection protocols.
output:
  papaja::apa6_word:
    keep_tex: yes
keywords: "Deception; Interoception; Theory of Mind; Poygraph; Lying Ability"
wordcount: '3720'
bibliography: ["references.bib"]
floatsintext: yes
figsintext: yes
figurelist: no
tablelist: no
footnotelist: no
linenumbers: yes
mask: yes
draft: no
documentclass: apa6
classoption: man
header-includes:
- \usepackage[labelfont=bf, font={color=gray,small}]{caption}
- \usepackage{float}
- \usepackage[document]{ragged2e}
csl: utils/apa.csl
---
